\sysfull

### 2017-08-11 paper/evaluation.tex
small fixes
evaluation section corrections

We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used for continuous power experiments. We used internal timer for measuring run time. For harvested energy, we used the ThingMagic Astra-EX RFID antenna set at 30\,dBm transmission power. We used EDB~\cite{edb} to measure run time. Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler. Version 3.8.0 was used for both Clang and LLVM.

@@

The applications that were used in the evaluation are: (i) \textbf{bc}: counts bits in byte pseudo-random string using varying algorithms. Each result is compared for correctness; (ii) \textbf{cem}: emulates an embedded temperature sensing engine: measurements are randomly generated and are compressed using a LZW algorithm for  fixed entry dictionary and fixed byte block; (iii) \textbf{cuckoo}: implements Cuckoo filter---the program first stores a sequence of pseudo-random numbers, then queries the filter to recover it; (iv) \textbf{rsa}: encrypts predefined string with a predefined key; (v) \textbf{ar} performs nearest neighbour classification using randomly generated sequence of three-axes accelerometer data. This set of benchmarks is the same as used in~\cite{chain,alpaca}. In addition to Chain benchmarks new applications were implemented: (vi) \textbf{dft}: discreet Fourier transform, and (vii) \textbf{sort}: selection sort algorithm. All applications are compared in Table~\ref{table:compiler_result} in terms of source lines of code against Chain.

@@

Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile the DFT application since LLVM inserts mathematical built-in functions (i.e. \texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}. We compiled the apps on a 4-core virtual machine with 1\,GB memory with Arch Linux, running on Intel Core i5-6200U (2.30\,GHz) processor. The result are given in Table~\ref{table:compiler_result}. We used the {\tt time} linux command to measure the rough time. Most of the compile time was due to pointer aliasing, when calculating possible live ranges of variables. Apps such as RSA whose code size is large and uses a lot of variables, need longer compilation time due to a larger search space for pointer aliasing. Nonetheless, the compile time is orders of magnitude shorter than manual task splitting by a programmer (in  order of hours). For this reason we believe that \sys complier will be a major step in enabling computing community to enter the intermittent-power domain without big time investments.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coala_chain}
	\caption{Performance of \sys applications (divided manually into tasks) at fixed power and at multiple WISP to RFID reader antenna distances, \{20, 40, 60\}\,cm, compared against Chain (results are normalized).}
	\label{fig:coalescing}
\end{figure}

We evaluate \sys experimentally to see its performance compared against Chain. Result is given in Figure~\ref{fig:coalescing}. For each application we measured execution time (with and without task coalescing) at \{20, 40, 60\}\,cm distance between RFID antenna and WISP and when WISP was powered directly from FET. Each measurement was for a predefined duration (each application was looping continuously) each distance/application case. From the result we clearly see an enormous benefit of \sys, for applications composed of small tasks (bc, sort, cuckoo)---at all distances and at fixed power. As the individual task become larger (e.g. dft) or program uses large memory pages (e.g. cem, refer to Section~
\ref{sec:result_compiler_time} for further explanation) \sys incurs overhead from memory virtualization that cause its performance to be comparable (or in a single cases, cem, worse) to Chain.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/comp_user}
	\caption{\sys compiler code overhead measured in terms of execution time per application; WISP was powered continuously.}
	\label{fig:comp_user}
\end{figure}

In Table~\ref{table:compiler_result} we have assessed the overhead induced by the \sys compiler. It is, nonetheless, important to find out how much does this overhead translates to increased execution time for each \sys application. The result of this investigation is provided in Figure~\ref{fig:comp_user}. We conclude that in most of the cases \sys compiler code provided comparable execution time to

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coalescing}
	\caption{\sys coalescing performance measured per application at fixed power and various WISP to RFID antenna distances (\{20, 40, 60\}\,cm) assessed against non-coalescing \sys.}
	\label{fig:coalescing}
\end{figure}

@@

Then, we need to know how many tasks \sys can coalesce within a given execution scenario. We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and the amount of tasks coalesced. Result is given in Figure~\ref{fig:aveVirtuTaskSize}. We see clearly that \sys manages to coalesce more tasks as individual tasks are small. As the size of individual task increases, e.g., as in the case of dft application, the number of coalesced tasks is also small. This result clearly shows that for the benefit of coalescing and for code portability, initial code should be split by \emph{as small tasks as possible}. This will help \sys to find the best possible virtual task size to maximize its runtime.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
	\caption{Normalized page size overhead and page faults per benchmark.\todo{Figure placeholder---will be replaced with final one}{Amjad}}
	\label{fig:IPOSPerformance}
	\label{fig:page_size}
\end{figure}

Figure~\ref{fig:page_size} presents the result on the effect of page swap size on the execution time of the programs executed by \sys. Overhead has been measured in the MCU clock cycles with values provided by TI CCS. Result has been normalized per benchmark: for the overhead results---to the minimum obtained value and for page faults---to the third measured maximum value\todo{Check again}{Amjad}. A subset of applications has been selected for evaluation. Additionally, page faults have been measured. The result show clearly a bathtub-shape bar plot for each application, demonstrating that for each application there is always a page size that minimizes \emph{both} execution time and page swaps. For all applications (except for Sort) this value equals roughly to 64 bytes. The shape of each figure can be explained as follows. For small page sizes (<64) the execution time is dominated by the \emph{frequency} of page swaps (each swap takes non-zero time, as shown in Figure~\ref{fig:dmaTimeEnergy}). On the other hand, in the opposite case the execution time is dominated by the time of swapping of one large page from SRAM to FRAM. Another interesting observation is the non-symmetric shape of each bar plot. This is due the size of data manipulated by the program: minimum point of overhead equals roughly to the size of the variables manipulated by each benchmark. Worth noting is also large discrepancy between minimum and maximum overhead for each benchmark (e.g. for Cuckoo 8 (at 16 bytes) to 1 (at 64 bytes), BC: 1 (at 16 bytes) to $\approx$2 (at 1024 bytes)). \todo{Explain this phenomenon}{Amjad}

>>>>

We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used for continuous power experiments. We used internal timer for measuring run time. For harvested energy, we used the ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power. We used EDB~\cite{edb} to measure run time. Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler. Version 3.8.0 was used for both Clang and LLVM.

@@

The applications that were used in the evaluation are: (i) \textbf{bc}: counts bits in byte pseudo-random string using varying algorithms. Each result is compared for correctness; (ii) \textbf{cem}: emulates an embedded temperature sensing engine: measurements are randomly generated and are compressed using a LZW algorithm for  fixed entry dictionary and fixed byte block; (iii) \textbf{cuckoo}: implements Cuckoo filter---the program first stores a sequence of pseudo-random numbers, then queries the filter to recover it; (iv) \textbf{rsa}: encrypts predefined string with a predefined key; (v) \textbf{ar} performs nearest neighbour classification using randomly generated sequence of three-axes accelerometer data. This set of benchmarks is the same as used in~\cite{chain,alpaca}. In addition to Chain benchmarks new applications were implemented: (vi) \textbf{sort}: selection sort algorithm, and (vii) \textbf{dft}: discreet Fourier transform. All applications are compared in Table~\ref{table:compiler_result} in terms of source lines of code against Chain.

@@

Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile the dft application since LLVM inserts mathematical built-in functions (i.e. \texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}. We compiled the apps on a 4-core virtual machine with 1\,GB memory with Arch Linux, running on Intel Core i5-6200U (2.30\,GHz) processor. The result are given in Table~\ref{table:compiler_result}. We used the {\tt time} linux command to measure the rough time. Most of the compile time was due to pointer aliasing, when calculating possible live ranges of variables. Apps such as RSA whose code size is large and uses a lot of variables, need longer compilation time due to a larger search space for pointer aliasing. Nonetheless, the compile time is orders of magnitude shorter than manual task splitting by a programmer (in  order of hours). For this reason we believe that \sys complier will be a major step in enabling computing community to enter the intermittent-power domain without big time investments.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coala_chain_clang}
	\caption{Performance of \sys applications (divided manually into tasks) at fixed power (left-most bar per application) and at multiple WISP to RFID reader antenna distances, \{20, 40\}\,cm (center and right-most bar per application) compared against Chain (results are normalized).}
	\label{fig:runtime}
\end{figure}

We evaluate \sys experimentally and compare its performance against Chain. Result is given in Figure~\ref{fig:runtime}. For each application we measured execution time (with and without task coalescing) at \{20, 40\}\,cm distance between RFID antenna and WISP, and when WISP was powered directly from FET. Each measurement was for a predefined duration (each application was looping continuously) each distance/application case. From the result we clearly see an enormous benefit of \sys, for applications composed of small tasks (bc, sort, cuckoo)---at all distances and at fixed power. As the individual task become larger (e.g. dft) or program uses large memory pages (e.g. cem, refer to Section~
\ref{sec:result_compiler_time} for further explanation) \sys incurs overhead from memory virtualization that cause its performance to be comparable (or in a single cases, cem, worse) to Chain.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/comp_user}
	\caption{\sys compiler code overhead measured in terms of execution time per application (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
	\label{fig:comp_user}
\end{figure}

In Table~\ref{table:compiler_result} we have assessed the overhead induced by the \sys compiler. It is, nonetheless, important to find out how much does this overhead translates to increased execution time for each \sys application. The result of this investigation is provided in Figure~\ref{fig:comp_user}. We conclude that in most of the cases \sys compiler code provided comparable execution time to code split into tasks manually (except for ar and rsa applications).

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/coalescing}
	\caption{\sys coalescing performance measured per application (left-most bar per application) and at multiple WISP to RFID reader antenna distances, \{20, 40\}\,cm (center and right-most bar per application) assessed against non-coalescing \sys.}
	\label{fig:coalescing}
\end{figure}

@@

Then, we need to know how many tasks \sys can coalesce within a given execution scenario. We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and the amount of tasks coalesced. Result is given in Figure~\ref{fig:aveVirtuTaskSize}. We see clearly that \sys manages to coalesce more tasks as individual tasks are small. As the size of individual task increases, e.g., as in the case of dft application, the number of coalesced tasks is also small. This result clearly shows that for the benefit of coalescing and for code portability, initial code should be split by \emph{as small tasks as possible}. This will help \sys to find the best possible virtual task size to minimise its runtime.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
	\caption{Normalized page size overhead and page faults per benchmark.}
	\label{fig:IPOSPerformance}
	\label{fig:page_size}
\end{figure}

Figure~\ref{fig:page_size} presents the result on the effect of page swap size on the execution time of the programs executed by \sys. Overhead has been measured in the MCU clock cycles with values provided by TI CCS. Result has been normalized per benchmark: for the overhead results---to the minimum obtained value and for page faults---to the third measured maximum value. A subset of applications has been selected for evaluation. Additionally, page faults have been measured. The result show clearly a bathtub-shape bar plot for each application, demonstrating that for each application there is always a page size that minimizes \emph{both} execution time and page swaps. For all applications (except for Sort) this value equals roughly to 64 bytes. The shape of each figure can be explained as follows. For small page sizes (<64) the execution time is dominated by the \emph{frequency} of page swaps (each swap takes non-zero time, as shown in Figure~\ref{fig:dmaTimeEnergy}). On the other hand, in the opposite case the execution time is dominated by the time of swapping of one large page from SRAM to FRAM. Another interesting observation is the non-symmetric shape of each bar plot. This is due the size of data manipulated by the program: minimum point of overhead equals roughly to the size of the variables manipulated by each benchmark. Worth noting is also large discrepancy between minimum and maximum overhead for each benchmark (e.g. for Cuckoo 8 (at 16 bytes) to 1 (at 64 bytes), BC: 1 (at 16 bytes) to $\approx$2 (at 1024 bytes)).

### 2017-08-11 paper/paging.tex
evaluation section corrections

\begin{algorithm}[t]
	\caption{\texttt{PageFault(tag)} pseudo-code}
	\label{algo:pagefault}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
		\If {\texttt{isDirty(CrntPagTag)} }	\Comment{Check if the active page is modified}
		\State \texttt{pageCommit(pagesTemp)} \Comment{Phase I commit the active page}
		\EndIf
		\If {\texttt{isDirty(tag)}} \Comment{Check if the requested page has been modified}
		\State {\texttt{getPage(tag,pagesTemp)}=false} \Comment{Get page from the temporary buffer}
		\Else
		\State \texttt{getPage(tag,pagesOrg)}=false \Comment{Get page from the original location}
		\EndIf 
		\State \texttt{CrntPagTag} = \texttt{tag} \Comment{Update tag}
	\end{algorithmic}
\end{algorithm}

>>>>

\begin{algorithm}[t]
	\caption{\texttt{PageFault(tag)} pseudo-code}
	\label{algo:pagefault}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
		\If {\texttt{isDirty(CrntPagTag)} }	\Comment{Check if the active page is modified}
		\State \texttt{pageCommit(pagesTemp)} \Comment{Phase I commit the active page}
		\EndIf
		\If {\texttt{isDirty(tag)}} \Comment{Check if the requested page has been modified}
		\State {\texttt{getPage(tag,pagesTemp)}} \Comment{Get page from the temporary buffer}
		\Else
		\State \texttt{getPage(tag,pagesOrg)} \Comment{Get page from the original location}
		\EndIf 
		\State \texttt{CrntPagTag} = \texttt{tag} \Comment{Update tag}
	\end{algorithmic}
\end{algorithm}
